---
title: "COX PH models- Concepts and implementation"
output: html_notebook
---



### Reading the data


```{r}
library("survival")
data(stanford2)
head(stanford2)

```
### Making a factor variable related to the Age > 40

```{r}

heart <- subset(stanford2, t5 != "NA")
Over40 <- ifelse(heart$age > 40,1,0)
heart <- cbind(heart, Over40)
heart$Over40 <- as.factor(heart$Over40)

```

### Making a factor variable related to T5 mismatchlevel 0,1,2


```{r}
len <- length(heart$t5)

misMatchLevel <- rep(0,len)

for (i in 1:len){
  
  if (heart$t5[i] > max(heart$t5)/3*2){
    
    misMatchLevel[i] <- 2
    
  }else if(heart$t5[i] > max(heart$t5)/3) {
    
    misMatchLevel[i] <- 1
  }
  
}

heart <- cbind(heart, misMatchLevel)

heart$misMatchLevel <- as.factor(heart$misMatchLevel)

```

### summary of the data

```{r}
summary(heart)
```
### Cox model for the Over40 and the mismatch level

```{r}
cox.mod <- coxph(Surv(heart$time, heart$status)~heart$Over40 + heart$misMatchLevel)
summary(cox.mod)
```
At a given instant in time, someone who is over 40 is 1.7249 time as as likely to die as someone who is under 40 adjusting for mismatch level


At a given instant in time, someone who is over 40 is 72.49 % time as as likely to die as someone who is under 40 adjusting for mismatch level

### Cox model for the Over40


```{r}
cox.mod2 <- coxph(Surv(heart$time, heart$status)~ heart$Over40)
```

### Analysing model differences


```{r}
anova(cox.mod2, cox.mod, test = "LRT")
```
When P chi is too high adding a coovarible is not statistically significant

### Cheking another model.

```{r}
cox.num <- coxph(Surv(heart$time, heart$status)~ heart$age +heart$t5)
summary(cox.num)
```
At a given instant in time, the probability of dying for someone who is 1 year older is 3% higher than someone who is 1 year younger adjusting for t5 score


### Checking the COX PH assumptions

#### Chekign linearity (For the model that used NUM X's)
#### usign martingale residuales

```{r}

plot(predict(cox.num), residuals(cox.num, type="martingale"),
     xlab= "Fitted values", ylab ="Martingale residuals",
     main="residual Plot", las = 1)
### add a line ax y = residual = 0
abline(h=0)

# fir a smoother thru the points

lines(smooth.spline(predict(cox.num), residuals(cox.num, type="martingale")), col = "red")
```


```{r}
### and checking linearity using deviance residuals

plot(predict(cox.num), residuals(cox.num, type="deviance"))
abline(h=0)

lines(smooth.spline(predict(cox.num), residuals(cox.num, type="deviance")), col = "red")
```

appears to be a non-linear... we can address this categorizing the non linear variable, transforming the variable or any other method, stratifying the variable or making it time dependent


#### 2. Proportional Hazard

Ho: hazards are prop Ha: hazards are not prop
will return test for each X,a nd for overall model

```{r}
cox.zph(cox.num)
```
All are larger than 5% so probably we should reject all of them


```{r}
par(mfrow=c(2,1))
plot(cox.zph(cox.num))
```
```{r}
par(mfrow=c(1,1))
plot(cox.zph(cox.num)[1])
abline(h=0, col=2)
```
```{r}
par(mfrow=c(1,1))
plot(cox.zph(cox.num)[2])
abline(h=0, col=2)
```
The red line fit a lot the CI most of th time, so we can say that fits the proportional hazard model